---
title: "Identify year sequences"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

df <- data.frame(year = c(1990:1995, 2000:2005), 
                 seq_id = c(rep(1, 6), rep(2, 6)))

df <- data.frame(year = c(1990:1995, 2000:2003),
                 seq_id = c(rep(1, 6), rep(2, 4)))

df <- data.frame(year = c(1990:1995, 2000, 2003:2005),
                 seq_id = c(rep(1, 6), 2, rep(3, 3)))

code_year_sequences <- function(x) {
  seq_vec <- vector("integer", length = length(x))
  seq_vec[1] <- 1
  for (i in 2:length(x)) {
    seq_vec[i] <- ifelse(x[i] - x[i-1]==1, seq_vec[i-1], seq_vec[i-1] + 1)
  }
  seq_vec
}



```

```{r}
library("lubridate")
library("stringr")
library("magrittr")
```


## Date helpers

### Last day of the month

```{r}
last_day_month <- function(x) {
  stopifnot(is.Date(x))
  x <- x %m+% months(1)
  day(x) <- 1
  day(x - 1)
}

```

### Date to ISO week and vice versa

Given an ISO week date like "2017-W01" with year and [ISO week](https://en.wikipedia.org/wiki/ISO_week_date), convert to a year-month-day date corresponding to the Thursday of that week. 

ISO weeks start on Monday and the first ISO week for a year is the first 4-day week in a year, which corresponds to the first Thursday in a year. As a result, the first and last ISO weeks of a year can include dates that fall in the previous and subsequent, respectively, calendar year. 

```{r}
isoweek2date <- function(x) {
  # input should be like "2014-W01"
  # output will be a date indexed to Thursday
  valid <- stringr::str_detect(x, "^[0-9]{4}-W[0-9]{2}$")
  stopifnot(all(valid))
  yy      <- substr(x, 1, 4) %>% as.integer()
  isoweek <- substr(x, 7, 8) %>% as.integer()
  
  first <- ymd(sprintf("%s-01-01", yy))
  # how many days to get to first Thursday of year?
  wday_first <- wday(first) - 1
  shift <- ifelse(wday_first > 4, 
                  11 - wday_first, 
                  4 - wday_first)
  
  out <- first + (isoweek - 1) * 7 + shift
  out
}
```

```{r}
test_dates <- data.frame(
  datestr = c("2015-W01", "2016-W01", "2017-W01", "2017-W02"),
  correct = c("2015-01-01", "2016-01-07", "2017-01-05", "2017-01-12"))
test_dates$date <- as.character(isoweek2date(test_dates$datestr))
test_dates

test_dates <- data.frame(
  datestr = c("2014-W52", "2015-W52", "2015-W53", "2016-W52"),
  correct = c("2014-12-25", "2015-12-24", "2015-12-31", "2016-12-29"))
test_dates$date <- as.character(isoweek2date(test_dates$datestr))
test_dates
```

The dates output by this function all fall on Thursday. To use a different weekday for indexing the weeks, add or subtract the difference in days between the desired index week day and Thursday, where Monday = 1 and Thursday = 4. E.g. to get weeks indexed by Monday, add $1 - 4 = -3$ to the output dates. 

```{r}
test_dates %>% isoweek2date() %>% add(-3) %>% wday(label = TRUE)
```


Convert a year-month-day date to corresponding year + ISO week

```{r}
date2isoweek <- function(x) {
  x <- as.Date(x, format = "%Y-%m-%d")
  out <- vector("character", length = length(x))
  x_good <- x[!is.na(x)]
  out[!is.na(x)] <- sprintf("%s-W%02s", 
                            lubridate::isoyear(x_good), 
                            lubridate::isoweek(x_good))
  out[is.na(x)]  <- NA_character_
  out
}

test_dates <- c("2015-01-01", "2015-01-02", "2017-01-05")
data.frame(date2 = test_dates, date1 = date2isoweek(test_dates))

test_dates <- c("2015-12-24", "2015-12-31", "2016-12-29")
data.frame(date2 = test_dates, date1 = date2isoweek(test_dates))
```



### Date ceiling and floor

See `?lubridate::floor_date`. 


